---
title: "Task 1"
author: "Group F, Ksenia Petrova, Nguyet Dao"
date: '2022-10-23'
output:
  word_document: default
  pdf_document: default
---

#### Topic

##### Research question:

Is there connection between total annual cost of electricity and total annual sales of the company?

##### Research hypothesis:

total annual cost of electricity is related to total annual sales.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, purl=TRUE}
data <- read.csv('https://raw.githubusercontent.com/adam-54/class/main/common/beeps_c.csv')

# subsetting by country
country_data <- subset(data, country =='Portugal')
```

#### 1. Target variable - total annual sales for all products 

```{r echo=FALSE, purl=TRUE}
d2 <- country_data$d2
```

##### Checking for missing data

```{r echo=FALSE}
x1 <- length(d2)
x2 <- sum(d2 == -9)
sprintf("Total number of values is %d", x1)
sprintf("Number of missing values is %d", x2)
```

##### Summary statistics

```{r echo=FALSE}
d2 <- replace(d2, d2 == -9, NA)
summary(d2)
x3 <- var(as.numeric(d2), na.rm = TRUE)
x4 <- sd(as.numeric(d2), na.rm = TRUE)
sprintf("Variance is equal to %f", x3)
sprintf("Standard deviation is equal to %f", x4)
```

##### Original target distribution

```{r echo=FALSE}
hist(as.numeric(country_data$d2), col = 'blue',
     xlab = 'Values',
     main = paste("Original target distribution"))
```

Target variable is skewed and requires transformation.

##### Imputation and log transformation

Share of missing values isn't too high, still they were imputed with the mean value. Log transformation was applied to data to conform to normality.

```{r echo=FALSE}
d2[is.na(d2)] <- mean(as.numeric(d2), na.rm = TRUE)
#d2 <- replace(d2, max(d2), mean(as.numeric(d2)))
d2 <- log10(d2)
```

##### Boxplot

```{r echo=FALSE}
boxplot(d2)
```

##### Histogram and distribution

```{r echo=FALSE}
library(ggplot2)
p11 = ggplot(as.data.frame(d2), aes(x=d2)) +
  geom_histogram(aes(y = ..density..), binwidth=.25, colour="black", fill="white") +
  xlab('Values') +
  stat_function(fun = dnorm, lwd = 2, col = 'red', 
                args = list(mean = mean(d2), sd = sd(d2)))
p11
```

As we can see from the histogram, target variable is distributed normally.

#### 2. Categorical predictor - establishment's main activity and product

| 1             | 2            | 3          | 4            | 5                   | 6                |
|--------------|-------------|----------|----------|-------------|----------------|
| Manufacturing | Retail trade | Whole sale | Construction | Hotel or restaraunt | Provide services |

\% of values in each category

```{r echo=FALSE}
# frequency
categories <- country_data$d1a1a
value_freq <- table(categories, useNA = 'ifany')

# percent
value_percent <- round(100 * (value_freq/length(categories)), 2)
value_percent

```

##### Boxplot by groups

```{r echo=FALSE}
dd <- cbind(d2, categories)
boxplot(dd[, 1] ~ dd[, 2],
        xlab = 'product categories',
        ylab = 'values')
```

##### Histogram of categorical predictor

```{r echo=FALSE}
hist(categories, 
     main = 'Categorical predictor',
     xlab = 'Product category')

```

##### Counting amount of NAs in each product category

```{r echo=FALSE}
library(dplyr)
df1 <- cbind(country_data$d2, country_data$d1a1a)
mydf <- data.frame(values = country_data$d2,
                  categories = country_data$d1a1a)
df2 <- replace(mydf, mydf == -9, NA)
na_count <- df2 %>% filter(is.na(values)) %>% group_by(categories) %>% summarise(n = n())
na_count
```

#### 3. Continuous predictor

```{r echo=FALSE}
n2b <- country_data$n2b
```

##### Checking missing data

```{r echo=FALSE}
x4 <- length(n2b)
x5 <- sum(n2b == -9)
sprintf("Total number of values is %d", x4)
sprintf("Number of missing values is %d", x5)
```

##### Imputation and log transformation

Missing data was imputed with the mean value. Data was log transformed also to better fit to target variable.

```{r echo=FALSE}
n2b <- replace(n2b, n2b == -9, mean(as.numeric(n2b)))
n2b <- log10(as.numeric(n2b))
```

##### Scatterplot of target and predictor

```{r echo=FALSE}
plot(d2, n2b, col = 'blue',
     xlab = 'Total annual sales',
     ylab = 'Total annual cost of electricity ')

```

##### How closely related two variables are (Pearson test)

```{r echo=FALSE}
res <- cor.test(country_data$d2, as.numeric(country_data$n2b), method = 'pearson')
cor_est <- round(as.numeric(res$estimate), 3)
sprintf("Correlation coefficient is equal to %f", cor_est)
```

Correlation between variables is rather high. It's expected that total annual cost of electricity and total annual sales are connected.


